---
title: "p8105_hw5_zx2528"
author: "Zizhi Xia_zx2528"
date: "2025-11-14"
output: github_document
---
# problem 1
```{r}
birthday_sim = function(n) {
  birthdays = sample(1:365, size = n, replace = TRUE)
  any(duplicated(birthdays))
}
```

```{r}
library(tidyverse)

set.seed(1234)

results =
  tibble(
    n = 2:50,
    prob = map_dbl(n, function(x) {
      mean(replicate(10000, birthday_sim(x)))
    })
  )
```

```{r}
results %>%
  ggplot(aes(x = n, y = prob)) +
  geom_line(size = 1.2) +
  geom_point(size = 1.8) +
  labs(
    title = "Probability That At Least Two People Share a Birthday",
    x = "Group size (n)",
    y = "Probability"
  ) +
  theme_minimal(base_size = 14)
```
We observe that the probability of shared birthdays increases rapidly as group size increases.
By the time the group size reaches 50, the probability is above 95%, showing that duplicates become almost guaranteed in moderately large groups despite having 365 possible birthdays.

# problem2
```{r}
library(broom)
library(tidyverse)

n <- 30
sigma <- 5
n_simulations <- 5000
alpha <- 0.05

mu_values <- c(0, 1, 2, 3, 4, 5, 6)

simulate_t_test <- function(mu, n, sigma) {

  x <- rnorm(n, mean = mu, sd = sigma)
  
  test_result <- t.test(x, mu = 0, alternative = "two.sided")
  
  tidy_result <- tidy(test_result)
  
  return(data.frame(
    mu_hat = tidy_result$estimate,
    p_value = tidy_result$p.value
  ))
}
```

```{r}
results <- data.frame()

for (mu in mu_values) {
  
  sim_results <- replicate(n_simulations, 
                           simulate_t_test(mu, n, sigma), 
                           simplify = FALSE)
  
  sim_df <- bind_rows(sim_results)
  sim_df$true_mu <- mu
  sim_df$rejected <- sim_df$p_value < alpha
  
  results <- bind_rows(results, sim_df)
}

power_summary <- results %>%
  group_by(true_mu) %>%
  summarise(
    power = mean(rejected),
    avg_estimate_all = mean(mu_hat),
    avg_estimate_rejected = mean(mu_hat[rejected]),
    n_rejected = sum(rejected)
  )

print(power_summary)
```

```{r}
plot1 <- ggplot(power_summary, aes(x = true_mu, y = power)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 3) +
  labs(
    title = "Relationship between Power and Effect Size",
    x = "True μ value",
    y = "Power (Proportion of H0 Rejections)"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1))

print(plot1)
```
Power increases as the true effect size μ increases.  
When μ = 0, the rejection rate is approximately 0.05, matching the Type I error rate.  
As μ becomes larger, the sample mean is more likely to differ substantially from 0, leading to higher power.
```{r}
plot2 <- ggplot(power_summary, aes(x = true_mu)) +
  geom_line(aes(y = avg_estimate_all, color = "All samples"), size = 1) +
  geom_point(aes(y = avg_estimate_all, color = "All samples"), size = 3) +
  geom_line(aes(y = avg_estimate_rejected, color = "Rejected samples only"), 
            size = 1, linetype = "dashed") +
  geom_point(aes(y = avg_estimate_rejected, color = "Rejected samples only"), size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "black") +
  labs(
    title = "Relationship between Average Estimate and True μ",
    x = "True μ value",
    y = "Average estimated μ̂ value",
    color = "Sample type"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("All samples" = "lightblue", 
                                 "Rejected samples only" = "pink"))
print(plot2)
```
Across all samples, the average estimated μ̂ closely matches the true value μ, confirming that the sample mean is an unbiased estimator.  
However, among *significant* results (p < 0.05), μ̂ is consistently larger than the true μ.  
This inflation is most pronounced when μ is small and power is low.

This simulation illustrates how effect size influences the performance of a one-sample t-test.  
Power increases with μ because larger effect sizes produce sample means further from zero, making it more likely to reject the null hypothesis.  

The sample mean itself is an unbiased estimator: across all replications, μ̂ closely matches the true μ.  
However, when conditioning on statistically significant findings, μ̂ becomes inflated—a manifestation of selection bias.  

Overall, these results demonstrate why statistically significant estimates tend to exaggerate effect sizes.

# problem3
```{r}
library(tidyverse)
library(broom)

homicide_df <- read_csv("data/homicide-data.csv")

homicide_df
```

```{r}
homicide_clean <- homicide_df %>%
  mutate(
    city_state = str_c(city, ", ", state),
    unsolved = disposition %in% c("Closed without arrest", "Open/No arrest")
  ) %>%
  group_by(city_state) %>%
  summarise(
    total = n(),
    unsolved = sum(unsolved)
  )

homicide_clean
```

```{r}
baltimore <- homicide_clean %>%
  filter(city_state == "Baltimore, MD")

baltimore_test <- prop.test(
  x = baltimore$unsolved,
  n = baltimore$total
)

baltimore_tidy <- broom::tidy(baltimore_test)

baltimore_tidy
```

```{r}
city_results <- homicide_clean %>%
  mutate(
    test = map2(unsolved, total, ~ prop.test(.x, .y)),
    tidy = map(test, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  select(city_state, estimate, conf.low, conf.high)

city_results
```

```{r}
city_results %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>%
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point(size = 2, color = "blue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  coord_flip() +
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimated proportion unsolved"
  ) +
  theme_minimal(base_size = 14)
```

The proportion of unsolved homicides varies widely across major U.S. cities.  
Baltimore, MD has one of the highest rates, with an estimated unsolved proportion of  
`r round(baltimore_tidy$estimate, 3)` and a 95% confidence interval of  
(`r round(baltimore_tidy$conf.low, 3)`, `r round(baltimore_tidy$conf.high, 3)`).  

Other cities exhibit substantially lower proportions, while some show even higher unsolved rates.  
These differences may reflect variations in investigative resources, community cooperation,  
case complexity, and police department practices.  

Confidence intervals are wider in cities with fewer total homicides, reflecting greater uncertainty,  
while larger cities provide more precise estimates.
